\begin{thebibliography}{20}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Abraham et~al.(2011)Abraham, Delling, Fiat, Goldberg, and
  Werneck]{AbrahamDFGW11}
I.~Abraham, D.~Delling, A.~Fiat, A.~V. Goldberg, and R.~F. Werneck.
\newblock {VC}-dimension and shortest path algorithms.
\newblock ICALP'11, 2011. 

\bibitem[Alon and Spencer(2008)]{AlonS08}
N.~Alon and J.~H. Spencer.
\newblock \emph{The Probabilistic Method}.
\newblock Interscience Series in Discrete Mathematics and Optimization. John
  Wiley {\&} Sons, Hoboken, NJ, USA, third edition, 2008.

\bibitem[Boucheron et~al.(2005)Boucheron, Bousquet, and Lugosi]{BoucheronBL05}
S.~Boucheron, O.~Bousquet, and G.~Lugosi.
\newblock Theory of classification : A survey of some recent advances.
\newblock \emph{{ESAIM}: Probability and Statistics}, 9:\penalty0 323--375,
  2005.

\bibitem[Br\"{o}nnimann and Goodrich(1994)]{BronnimannG94}
H.~Br\"{o}nnimann and M.T. Goodrich.
\newblock Almost Optimal Set Covers in Finite VC-Dimension
\newblock SCG'94, 1994.

\bibitem[Cormode and Duffield(2014)]{CormodeD14}
G.~Cormode and N.~Duffield.
\newblock Sampling for Big Data: A tutorial. 
\newblock KDD'14, 2014.

\bibitem[Kearns and Schapire(1990)]{KearnsS90}
M.~J. Kearns and R.~E. Schapire.
\newblock Efficient distribution-free learning of probabilistic concepts.
\newblock FOCS'90, 1990.

\bibitem[Kleinberg et~al.(2008)Kleinberg, Sandler, and Slivkins]{KleinbergSS08}
J.~M. Kleinberg, M.~Sandler, and A.~Slivkins.
\newblock Network failure detection and graph connectivity.
\newblock \emph{SIAM J. Comput.}, 38\penalty0 (4):\penalty0 1330--1346, 2008.

\bibitem[Koltchinskii(2001)]{Koltchinskii01}
V.~Koltchinskii.
\newblock Rademacher penalties and structural risk minimization.
\newblock \emph{IEEE Trans.~Inf.~Theory}, 47\penalty0 (5):\penalty0 1902--1914,
  July 2001.

\bibitem[Kranakis et~al.(1997)Kranakis, Krizanc, Ruf, Urrutia, and
  Woeginger]{KranakisKRUW97}
E.~Kranakis, D.~Krizanc, B.~Ruf, J.~Urrutia, and G.~Woeginger.
\newblock The {VC}-dimension of set systems defined by graphs.
\newblock \emph{Discrete Applied Mathematics}, 77\penalty0 (3):\penalty0
  237--257, 1997.

\bibitem[Mohri et~al.(2012)Mohri, Rostamizadeh, and Talwalkar]{MohriRT12}
M.~Mohri, A.~Rostamizadeh, and A.~Talwalkar.
\newblock \emph{Foundations of machine learning}.
\newblock The MIT Press, 2012.

\bibitem[Papadimitriou and Yannakakis(1996)]{PapadimitriouY96}
C.~H. Papadimitriou and M.~Yannakakis.
\newblock On limited nondeterminism and the complexity of the {V-C} dimension.
\newblock \emph{J. Comput. Syst. Sci.}, 53\penalty0 (2):\penalty0 161--170,
  1996.

\bibitem[Pollard(1984)]{Pollard1984}
D.~Pollard.
\newblock \emph{Convergence of stochastic processes}.
\newblock Springer, 1984.

\bibitem[Riondato(2014)]{Riondato14}
M.~Riondato.
\newblock Sampling-based data mining algorithms: Modern techniques and case
  studies.
\newblock ECML PKDD'14, 2014.

\bibitem[Riondato and Kornaropoulos(2014)]{RiondatoK14}
M.~Riondato and E.~M. Kornaropoulos.
\newblock Fast approximation of betweenness centrality through sampling.
\newblock WSDM'14, 2014.

\bibitem[Riondato and Upfal()]{RiondatoU15-ext}
M.~Riondato and E.~Upfal.
\newblock Mining frequent itemsets through progressive sampling with rademacher
  averages.
\newblock Extended Version.
\newblock URL \url{http://cs.brown.edu/%7Ematteo/papers/progrsamplfi-ext.pdf}.

\bibitem[Riondato and Upfal(2014)]{RiondatoU14}
M.~Riondato and E.~Upfal.
\newblock Efficient discovery of association rules and frequent itemsets
  through sampling with tight performance guarantees.
\newblock \emph{ACM Trans. Knowl. Disc. from Data}, 8\penalty0 (4):\penalty0
  20, 2014.

\bibitem[Shalev-Shwartz and Ben-David(2014)]{ShalevSBD14}
S.~Shalev-Shwartz and S.~Ben-David.
\newblock \emph{Understanding Machine Learning: From Theory to Algorithms}.
\newblock Cambridge University Press, 2014.

\bibitem[Shao et~al.(2000)Shao, Cherkassky, and Li]{ShaoCL00}
X.~Shao, V.~Cherkassky, and W.~Li.
\newblock Measuring the {VC}-dimension using optimized experimental design.
\newblock \emph{Neural Computation}, 12\penalty0 (8):\penalty0 1969--1986,
  2000.

\bibitem[Vapnik(1999)]{Vapnik99}
V.~N. Vapnik.
\newblock \emph{The Nature of Statistical Learning Theory}.
\newblock Statistics for engineering and information science. Springer-Verlag,
  New York, NY, USA, 1999.

\bibitem[Vapnik and Chervonenkis(1971)]{VapnikC71}
V.~N. Vapnik and A.~J. Chervonenkis.
\newblock On the uniform convergence of relative frequencies of events to their
  probabilities.
\newblock \emph{Theory of Probability and its Applications}, 16\penalty0
  (2):\penalty0 264--280, 1971.

\bibitem[Vapnik et~al.(1994)Vapnik, Levin, and Le~Cun]{VapnikLLC94}
V.~N. Vapnik, E.~Levin, and Y.~Le~Cun.
\newblock Measuring the {VC}-dimension of a learning machine.
\newblock \emph{Neural Computation}, 6:\penalty0 851--876, 1994.

\end{thebibliography}
